---
title: "CH 10.1 - 10.6: Multiple Predictors"
output: pdf_document
---

\renewcommand{\vec}[1]{\mathbf{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 3, fig.width = 5, fig.align = 'center')
library(tidyverse) 
library(gridExtra)
library(rstanarm)
library(arm)
set.seed(10052020)
```

### Regression with Multiple Predictors

```{r, echo = F, message = F}
beer <- read_csv('http://math.montana.edu/ahoegh/Data/Brazil_cerveja.csv')
```

#### Binary Predictor

With a single binary predictor, using the beer dataset, the model can be written as

$$y = \beta_0 + \beta_1 x_{weekend = 1} + \epsilon,$$
where:

- $y$ is the beer consumption, 
\vfill

- $\beta_0$ is the consumption on a weekday, 
\vfill

- $\beta_1$ is the difference in consumption between a weekend day and a weekend,  
\vfill

- $x_{weekend = 1}$ is an indicator function for whether the observation is a weekend. 

\vfill


\vfill
```{r}
stan_glm(consumed ~ weekend, data = beer, refresh = 0)
```

\vfill

\newpage

Alternatively, the cell means model can be constructed, where

$$y = \beta_1 x_{weekend = 0} + \beta_2 x_{weekend = 1} + \epsilon,$$
in this case 

- $\beta_1$ is the mean consumption on a weekday \vfill
- $\beta_2$ is the mean consumption on a weekend. \vfill


\vfill

```{r}
beer <- beer %>% mutate(weekend = factor(weekend))
stan_cm <- stan_glm(consumed ~ weekend - 1, data = beer, refresh = 0)
stan_cm
```

\vfill

\vfill
```{r}
as.data.frame(stan_cm) %>% mutate(contrast = weekend1 - weekend0) %>%
  summarise(median_diff = median(contrast), lower_interval = quantile(contrast, probs = .025), 
            upper_interval = quantile(contrast, probs = .975))
```


\newpage
#### Binary Predictor + Continuous Predictor

Now consider jointly considering both weekend/weekday and maximum temperature. The model can now be written as

$$y = \beta_0 + \beta_1 x_{weekend = 1} + \beta_2 x_{tmp} + \epsilon,$$
where:

- $y$ is the beer consumption, 
\vfill

- $\beta_0$ 

\vfill

- $\beta_1$ 
\vfill

- $\beta_2$ 

\vfill

```{r}
ml_regression <- beer %>% stan_glm(consumed ~ weekend + max_tmp, data = ., refresh = 0)
ml_regression
```

\newpage

#### Coefficient Interpretation

- When interpreting coefficients in a multiple regression model, it is important to understand that these values control for other predictors in the model! 

\vfill

- Sometimes we cannot necessarily hold all other predictors constant in a model. A simple example would be $y = \beta_0 + \beta_1 x + \beta_2 x^2 + \epsilon$

\vfill

The textbook puts an emphasis on differentiating predictive and counterfactual interpretations:

- The predictive interpretation focuses on how the outcome differs, on average, when _comparing_ two groups of items that differ by 1 unit (and all other predictors are the same). 

\vfill
- The counterfactual interpretation focuses on how the outcome would differ with an individual, rather than between individuals. 

\vfill

The counterfactual interpretation should be reserved for a situation where causal inferences are reasonable, such as a completely randomized experimental design.

\vfill

It is easy to get careless with wording and say things like "a change in temperature is associated with a change in consumption," but 

\vfill

\newpage

#### Interaction

The model what we have fit, results in two parallel lines.
```{r, fig.width = 6, fig.height = 4}
beer %>% ggplot(aes(y = consumed, x = max_tmp, color = weekend)) +
  geom_point() + 
  geom_abline(intercept = as.numeric(ml_regression$coefficients[1]),
              slope = as.numeric(ml_regression$coefficients[3]), color = "#E69F00") + 
    geom_abline(intercept = as.numeric(ml_regression$coefficients[1] + ml_regression$coefficients[2]),
              slope = as.numeric(ml_regression$coefficients[3]), color = "#56B4E9") + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  theme_minimal() + 
  xlab("Maximum Temperature (C)") + 
  ylab("Liters of Beer Consumed") + 
  ggtitle('Beer consumption in Sao Paolo, Bazil') +
  labs(caption = 'Lines represent model fit without interaction')
```

In this situation, the assumption of an additive model (parallel lines) seems reasonable. However, in many situations we'd expect that the relationship between a continuous covariate (e.g. an extra degree of maximum temperature) could depend upon another covariate (e.g. day of week). 

\vfill

With our data, an interaction would mean that the two lines _are not_ parallel.

\vfill

The next figure allows some flexibility to fit non-parallel (and non-linear) functional relationships.


```{r, eval = F}
beer %>% ggplot(aes(y = consumed, x = max_tmp, color = weekend)) + geom_point() + geom_smooth(formula ='y~x', method = 'loess', se = F) + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) + 
    theme_minimal() + 
  xlab("Maximum Temperature (C)") + 
  ylab("Liters of Beer Consumed") + 
  ggtitle('Beer consumption in Sao Paolo, Bazil') +
  labs(caption = 'Lines represent loess fit')
```
\vfill
\newpage

```{r, eval = T, echo = F,  fig.width = 6, fig.height = 4}
beer %>% ggplot(aes(y = consumed, x = max_tmp, color = weekend)) + geom_point() + geom_smooth(formula ='y~x', method = 'loess', se = F) + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) + 
  theme_minimal() + 
  xlab("Maximum Temperature (C)") + 
  ylab("Liters of Beer Consumed") + 
  ggtitle('Beer consumption in Sao Paolo, Bazil') +
  labs(caption = 'Lines represent loess fit')
```

This figure doesn't suggest a non-additive relationship, but nevertheless, let's explore the interaction model.

\vfill


$$y = \beta_0 + \beta_1 x_{weekend = 1} + \beta_2 x_{tmp} + \beta_3 x_{weekend = 1} x_{tmp} +  \epsilon,$$

- $\beta_0$ 

\vfill

- $\beta_1$ 
\vfill

- $\beta_2$ 
\vfill

- $\beta_3$ 
\vfill
\newpage

```{r}
stan_glm(consumed ~ max_tmp * weekend, data = beer, refresh = 0)
```
\vfill

Interaction coefficients are more easily interpreted when the the continuous variables are centered or standardized.

\vfill

R automatically creates indicator variables for categorical data (that are stored as factors). 

\vfill

```{r}
model.matrix(consumed ~ weekend, data = beer) %>% head(3)
model.matrix(consumed ~ weekend - 1, data = beer) %>% head(3)
```

\vfill

To change the reference level, it is necessary to either directly specify the levels of the factor or reorder them (see `forcats`)
```{r}
beer %>% mutate(weekend_fact = factor(weekend, levels = c('1', '0'))) %>%
  lm(consumed ~ weekend_fact, data = .) %>% display()
```

\vfill
\newpage

#### Computational Lab

The purpose of this activity is to better understand an interaction model.



1. Simulate fake data that has an interaction.
\vfill


2. Visualize the interaction.

\vfill

3. Fit interaction model.

\vfill